HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-07-15 scnn.tex
figures: textwidth => linewidth

A second direction has been followed in~\citet{boomsma2017spherical}. The idea is
to use the traditional 2-dimensional convolution on an irregular grid defined on the
sphere. In the paper, two grids are used: the equirectangular grid and the ``cubed
sphere'' grid defined by~\citet{ronchi1996cubed}. \nati{Do we need a plot of the equirectangular and cubed sphere sampling?} Because this
approach is based on the traditional 2-dimensional convolution, it is very likely to be the
most computationally efficient one. However, it suffers that it can only be used using very
specific grid sampling sets which are not including HEALPix. Furthermore, the
grid requirement makes it impossible for the convolution to capture the
spherical properties of the domain, i.e., the "cubed sphere” sampling is by
definition adapted to a cube and not to a sphere.
\todo{border effects? \nati{@mdeff what are you thinking about here?}}
% \nati{Michael: Under some very
% specific hypothesis, this second approach is a particular case of our method,
% i.e. 1) our method with a stupid sampling, 2) assumption that graph convolution
% on a grid == 2d convolution. Do you think we should mention that? I think we
% should not.}
% I think we should!

@@

% mdeff: Why not $1/d$? See Pascal's paper
% nati: The question is: is $1/d$ also good when we have 7/8 neighboors instead of 4? Practically, we can also use $1/d$, but that means changing the code and re-runing the experiment. We probably do not have to change any parameter though... @mdeff: do you think we should try this for this paper? Or we keep that for the next one.
% mdeff: I tried and the eigenvectors and eigenvalues are mostly the same. To me, there is no point in trying for trying. We should rather come up with a definition of distance that will make the eigenvectors closer to the real spherical harmonics.

\begin{figure}
	\centering
	% \includegraphics[width=0.45\textwidth]{figures/healpix_graph_4.pdf}
	\includegraphics[width=0.8\linewidth]{figures/half_graph_4.pdf}
	\caption[]{Graph built from an HEALPix pixelization of half the sphere ($N_{side} = 4$). By construction, each node has 8 neighbors, except the highlighted ones which have only 7.\footnotemark[6]}
	\label{fig:healpix_graph_4}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{figures/gaussian_filters_sphere.pdf}
	\caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
	Top: Diffusion of a unit of heat for different times $t$ using the graph.
	Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
	Relative difference between graph convolution and spherical smoothing: $10.4$\%, $5.7$\%, $4.8$\%, $3.8$\% }
	\label{fig:gaussian_filters_comparizon}
\end{figure*}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/pooling.pdf}
	\caption{Two levels of coarsening and pooling: groups of 4 cells are merged into one, then the data on them is summarized in one. The coarsest cell covers $1/12$ of the sphere.}
	\label{fig:pooling}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.45\textwidth]{figures/psd_sigma3.pdf}
\caption{Power spectrum densities of the noiseless maps. The maps have been smoothed with a Gaussian kernel of radius $3$ arcmins to remove high frequencies. The cosmological models could otherwise have been distinguished by looking at the PSDs.}
\label{fig:psd_sigma3}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.48\textwidth]{figures/smooth_map_class_1.pdf}
\includegraphics[width=0.48\textwidth]{figures/smooth_map_class_2.pdf}
\caption{Example maps from two classes to be discriminated. Left: model 1 with $\Omega_m=0.31$ and $\sigma_8=0.82$. Right: model 2 with $\Omega_m=0.26$ and $\sigma_8=0.91$.
The initial conditions for both simulations were the same, so differences only arise due to different cosmological parameters. \nati{Tomek: Is this correct?}}
\label{fig:map_sample}
\end{figure*}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.45\textwidth]{figures/part_sphere.pdf}
\caption{3 subparts of the sphere with different sizes. Blue: $o=1$. Green: order $o=2$. Yellow: order $o=4$.}
\label{fig:subpart_sphere}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.32\textwidth]{figures/result_order1.pdf}
\includegraphics[width=0.32\textwidth]{figures/result_order2.pdf}
\includegraphics[width=0.32\textwidth]{figures/result_order4.pdf}
\caption{Classification errors for the 3 different problems.}
\label{fig:results}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.45\textwidth]{figures/section_filter_last.pdf}
\includegraphics[width=0.45\textwidth]{figures/gnonomic_filter_last.pdf}
\caption{24 learned filters selected from the last spherical convolutional layer. Left: section. Right: gnonomic view.}
\label{fig:learned_filter}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.95\textwidth]{figures/gaussian_filters_spectral.pdf}
\includegraphics[width=0.95\textwidth]{figures/gaussian_filters_gnomonic.pdf}
\includegraphics[width=0.95\textwidth]{figures/gaussian_filters_section.pdf}
\caption{3 different visualizations of the convolution kernel $K_t(x)=e^{-\tau t x}$.
Top: graph spectral domain.
Middle: gnomonic projection.
Bottom: section plot.}
\label{fig:gaussian_filters_visualization}
\end{figure*}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.3\textwidth]{figures/index_plotting_order20_nside16.pdf}
\caption{Indexes selected for the section ploting of \figref{gaussian_filters_visualization} middle. The delta was placed on the yellow node.}
\label{fig:index_section}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.45\textwidth]{figures/hist_error_order2_noise1_5.pdf}
\caption{Evolution curve of the error with respect to the number of training samples for histogram features and a linear classifier (setting: order 2, relative noise level: 1.5). The validation error might be slightly below the training error because we use it for cross-validation.}
\label{fig:hist_error_evolution}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.45\textwidth]{figures/border_effects.pdf}
\caption{Graph convolution of a Kroneker placed at two different spots on the graph. The graph build from $1/12$ of the sphere and an Nside of 16. When the Kroneker is placed on the left corner, the resulting convolution is not circular anymore due to border effect. The graph setting assumes reflective border conditions.}
\label{fig:border_effects}
\end{figure}

>>>>

A second direction has been followed in~\citet{boomsma2017spherical}. The idea is
to use the traditional 2-dimensional convolution on an irregular grid defined on the
sphere. In the paper, two grids are used: the equirectangular grid and the ``cubed
sphere'' grid defined by~\citet{ronchi1996cubed}. \nati{Do we need a plot of the equirectangular and cubed sphere sampling?} Because this
approach is based on the traditional 2-dimensional convolution, it is very likely to be the
most computationally efficient one. However, it suffers that it can only be used using very
specific grid sampling sets which are not including HEALPix. Furthermore, the
grid requirement makes it impossible for the convolution to capture the
spherical properties of the domain, i.e., the ``cubed sphere'' sampling is by
definition adapted to a cube and not to a sphere.
\todo{border effects? \nati{@mdeff what are you thinking about here?}}
% \nati{Michael: Under some very
% specific hypothesis, this second approach is a particular case of our method,
% i.e. 1) our method with a stupid sampling, 2) assumption that graph convolution
% on a grid == 2d convolution. Do you think we should mention that? I think we
% should not.}
% I think we should!

@@

% mdeff: Why not $1/d$? See Pascal's paper
% nati: The question is: is $1/d$ also good when we have 7/8 neighbors instead of 4? Practically, we can also use $1/d$, but that means changing the code and re-runing the experiment. We probably do not have to change any parameter though... @mdeff: do you think we should try this for this paper? Or we keep that for the next one.
% mdeff: I tried and the eigenvectors and eigenvalues are mostly the same. To me, there is no point in trying for trying. We should rather come up with a definition of distance that will make the eigenvectors closer to the real spherical harmonics.

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{figures/half_graph_4.pdf}
	\caption[]{Graph built from an HEALPix pixelization of half the sphere ($N_{side} = 4$). By construction, each node has 8 neighbors, except the highlighted ones which have only 7.\footnotemark[6]}
	\label{fig:healpix_graph_4}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{figures/gaussian_filters_sphere.pdf}
	\caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
	Top: Diffusion of a unit of heat for different times $t$ using the graph.
	Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
	Relative difference between graph convolution and spherical smoothing: $10.4$\%, $5.7$\%, $4.8$\%, $3.8$\% }
	\label{fig:gaussian_filters_comparizon}
\end{figure*}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{figures/pooling.pdf}
	\caption{Two levels of coarsening and pooling: groups of 4 cells are merged into one, then the data on them is summarized in one. The coarsest cell covers $1/12$ of the sphere.}
	\label{fig:pooling}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/psd_sigma3.pdf}
\caption{Power spectrum densities of the noiseless maps. The maps have been smoothed with a Gaussian kernel of radius $3$ arcmins to remove high frequencies. The cosmological models could otherwise have been distinguished by looking at the PSDs.}
\label{fig:psd_sigma3}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.48\linewidth]{figures/smooth_map_class_1.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{figures/smooth_map_class_2.pdf}
\caption{Example maps from two classes to be discriminated. Left: model 1 with $\Omega_m=0.31$ and $\sigma_8=0.82$. Right: model 2 with $\Omega_m=0.26$ and $\sigma_8=0.91$.
The initial conditions for both simulations were the same, so differences only arise due to different cosmological parameters. \nati{Tomek: Is this correct?}}
\label{fig:map_sample}
\end{figure*}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/part_sphere.pdf}
\caption{Three parts of the sphere with different sizes. Blue: $o=1$. Green: order $o=2$. Yellow: order $o=4$.}
\label{fig:subpart_sphere}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.32\linewidth]{figures/result_order1.pdf}
\hfill
\includegraphics[width=0.32\linewidth]{figures/result_order2.pdf}
\hfill
\includegraphics[width=0.32\linewidth]{figures/result_order4.pdf}
\caption{Classification errors for the 3 different problems.}
\label{fig:results}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.47\linewidth]{figures/section_filter_last.pdf}
\hfill
\includegraphics[width=0.45\linewidth]{figures/gnonomic_filter_last.pdf}
\caption{24 learned filters selected from the last spherical convolutional layer. Left: section. Right: gnonomic view.}
\label{fig:learned_filter}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/gaussian_filters_spectral.pdf}
\includegraphics[width=\linewidth]{figures/gaussian_filters_gnomonic.pdf}
\includegraphics[width=\linewidth]{figures/gaussian_filters_section.pdf}
\caption{Three different visualizations of the convolution kernel $K_t(x)=e^{-\tau t x}$.
Top: graph spectral domain.
Middle: gnomonic projection.
Bottom: section plot.}
\label{fig:gaussian_filters_visualization}
\end{figure*}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.6\linewidth]{figures/index_plotting_order20_nside16.pdf}
\caption{Indexes selected for the section ploting of \figref{gaussian_filters_visualization} middle. The delta was placed on the yellow node.}
\label{fig:index_section}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/hist_error_order2_noise1_5.pdf}
\caption{Evolution curve of the error with respect to the number of training samples for histogram features and a linear classifier (setting: order 2, relative noise level: 1.5). The validation error might be slightly below the training error because we use it for cross-validation.}
\label{fig:hist_error_evolution}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/border_effects.pdf}
\caption{Graph convolution of a Kroneker placed at two different spots on the graph. The graph build from $1/12$ of the sphere and an Nside of 16. When the Kroneker is placed on the left corner, the resulting convolution is not circular anymore due to border effect. The graph setting assumes reflective border conditions.}
\label{fig:border_effects}
\end{figure}
